Analysis paper

### 2021-06-21 analysis/analysis.Rmd
update plots

```{r}
dailytruth_data %>%
  dplyr::filter(target_end_date >= min(forecast_dates$full_hub_period), 
                target_end_date <= max(forecast_dates$full_hub_period)) %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_point(size = 0.3, alpha = 0.4) + 
  geom_line(size = 0.5, alpha = 0.4) + 
  geom_line(data = truth_data, aes(y = true_value / 7)) + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                            max(forecast_dates$full_hub_period)))
  theme_minimal()

@@

```{r truth-vs-forecasts, fig.height=12, fig.width=15}
# create helper function for the plot ------------------------------------------
truth_and_pred_plot <- function(df, ttype = "case") {
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                   size = 1.7,
                   alpha = 1,
                   position = position_dodge(width = 4)) + 
    # geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
    #                size = 1.7,
    #                alpha = 0.1,
    #                position = position_dodge(width = 4)) +   
    geom_point(data = unfiltered_data[target_type == ttype], 
               aes(shape = classification)) + 
    geom_line(data = unfiltered_data[target_type == ttype]) + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                            max(forecast_dates$full_hub_period)))
  return(p)
}

@@

df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, forecast_date := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

@@

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_cases"))
  ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_cases)

@@

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_deaths"))
  ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_deaths)

@@

```{r scores-over-time, fig.height=10, fig.width=15}
# score forecasts --------------------------------------------------------------
# obtain scores and then do pairwise comparisons between models
summarise_by <- c("model", "horizon", "location_name", "target_type", 
                   "forecast_date", "target_end_date", "location_target", 
                   "classification")
scores_target_forecastdate <- eval_forecasts(
  data = unfiltered_data[model %in% regular_models],
  summarise_by = summarise_by
)

@@

# create helper function to plot forecasts -------------------------------------
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(shape = classification, colour = model)) + 
  theme_minimal() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  # geom_rect(aes(xmin = min(forecast_dates$christmas) + 14,
  #               xmax = max(forecast_dates$christmas) + 14,
  #               ymin = 0,
  #               ymax = Inf), fill = "lightgrey", alpha = 0.002) +
  # theme(panel.background = element_rect(fill = "#e1f0fe")) +
  # scale_y_log10() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

# create plots for all horizons ------------------------------------------------
for (i in 1:4) {
  scores_over_time_case <- scores_target_forecastdate[horizon == i & target_type == "case"] %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = colcases) + 
    labs(y = "WIS rel. to baseline", x = "")

@@

  scores_over_time_deaths <- scores_target_forecastdate[horizon == i & target_type == "death"] %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = coldeaths) + 
    annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
             xmax = max(forecast_dates$death_not_scored) + 14,
             ymin = 0,
             ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "WIS rel. to baseline", x = "Date")

@@

comparison_ensembles <- function(models_to_plot) {
  df <- unfiltered_data[model %in% models_to_plot &
                          quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, forecast_date := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

@@

```{r}
# only select every second forecast date to stratify
df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, forecast_date := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

>>>>

```{r}
dailytruth_data %>%
  dplyr::filter(target_end_date >= min(forecast_dates$full_hub_period), 
                target_end_date <= max(forecast_dates$full_hub_period)) %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_point(size = 0.3, alpha = 0.4) + 
  geom_line(size = 0.5, alpha = 0.4) + 
  geom_line(data = truth_data, aes(y = true_value / 7)) + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  labs(y = "Observed values", x = "Date") + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) - 28, 
                            max(forecast_dates$full_hub_period))) + 
  theme_minimal()

@@

```{r truth-vs-forecasts, fig.height=12, fig.width=15}
# create helper function for the plot ------------------------------------------
truth_and_pred_plot <- function(df, ttype = "case") {

  trend <- copy(epitrend)[, location_target := paste0(str_to_sentence(target_type), 
                                                    "s", " in ", location_name)]
  p <- df %>%
    dplyr::filter(target_type == ttype) %>%
    ggplot(aes(y = true_value, x = target_end_date)) + 
    geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                   size = 1.7,
                   alpha = 1,
                   position = position_dodge(width = 4.6)) + 
    # geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
    #                size = 1.7,
    #                alpha = 0.1,
    #                position = position_dodge(width = 4)) +   
    geom_point(data = trend[target_type == ttype],
               aes(shape = classification)) +
    # geom_point(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    geom_line(data = unfiltered_data[target_type == ttype & horizon == 2]) + 
    theme_minimal() + 
    facet_wrap(~ location_target, scales = "free", ncol = 2) + 
    theme(legend.position = "bottom") + 
    # theme(panel.background = element_rect(fill = "#e1f0fe")) + 
    scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                            max(forecast_dates$full_hub_period)))
  return(p)
}

@@

df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

@@

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_cases"))
  # ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_cases)

@@

  filename <- here("analysis", "plots", 
                   paste0("truth-and-", i, "-wk-ahead_deaths"))
  # ggsave(paste0(filename, ".png"), plot = truth_vs_forecast_deaths)

@@

```{r scores-over-time, fig.height=10, fig.width=15}
# score forecasts --------------------------------------------------------------
# obtain scores and then do pairwise comparisons between models
summarise_by <- c("model", "horizon", "location_name", "target_type", 
                   "forecast_date", "target_end_date", "location_target")

scores_target_forecastdate <- eval_forecasts(
  data = unfiltered_data[model %in% regular_models],
  summarise_by = summarise_by
)

@@

# create helper function to plot forecasts -------------------------------------
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(colour = model, shape = classification)) + 
  theme_minimal() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  # geom_rect(aes(xmin = min(forecast_dates$christmas) + 14,
  #               xmax = max(forecast_dates$christmas) + 14,
  #               ymin = 0,
  #               ymax = Inf), fill = "lightgrey", alpha = 0.002) +
  # theme(panel.background = element_rect(fill = "#e1f0fe")) +
  # scale_y_log10() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 5, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

# create plots for all horizons ------------------------------------------------
for (i in 1:4) {
  epi_tmp <- copy(epitrend)
  epi_tmp[, forecast_date := as.character(as.Date(forecast_date) + 7 * i)]
  epi_tmp <- epi_tmp[, .(location_name, forecast_date, classification, target_type)] %>% 
    unique()

  scores_over_time_case <- scores_target_forecastdate[horizon == i & target_type == "case"] %>%
    merge(epi_tmp) %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = colcases) + 
    labs(y = "WIS rel. to baseline", x = "")

@@

  scores_over_time_deaths <- scores_target_forecastdate[horizon == i & target_type == "death"] %>%
    merge(epi_tmp) %>%
    scores_over_time_plot(y = "interval_score") + 
    scale_color_manual(values = coldeaths) + 
    annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
             xmax = max(forecast_dates$death_not_scored) + 14,
             ymin = 0,
             ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "WIS rel. to baseline", x = "Date")

@@

comparison_ensembles <- function(models_to_plot) {
  df <- unfiltered_data[model %in% models_to_plot &
                          quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
    dcast(... ~ quantile, value.var = "prediction")
  df <- merge(
    df[, c("classification", "speed") := NULL], 
    copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
    by = c("target_end_date", "location_name", "target_type")
  )

@@

```{r}
# only select every second forecast date to stratify
df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.5, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, c("forecast_date", "true_value") := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)
